RNApyro

### 2012-10-05 Recomb/methods_RECOMB.tex
methods

%!TEX root = main_RECOMB.tex
\section{Methods}
\label{sec:methods}
Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
all possible secondary structures of length  $s$. We are interested in  
the probability of a given position being a specific nucleotides,
over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ (i.e.
$\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
We define a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
 desired probability, with the two following functions,
$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, as the 
equivalent of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular, knowing 
that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing outside,
including $i$ and $j$, $m$ mutations.
The Boltzmann weights are a combination of the base pairs stacking energy, 
using as values those of the NNDB~\cite{Turner2010}, and
the average isostericity with $\Omega$, as defined in~\cite{Stombaugh2009}.

>>>>

%!TEX root = main_RECOMB.tex
\section{Methods}
\label{sec:methods}
Let $\Omega$ be an un-gapped RNA alignment, $S$ its associated secondary structure, $s$ an RNA 
sequence and $m\geq 0$. $S$ is considered as  one derivation of the SCFG generating  
all  secondary structures of length  $|s|$. We are interested in  
the probability of a given position being a specific nucleotides,
over all sequences having at most $m$ mutations from $s$, under the SCFG derivation $S$ 
(formally 
$\mathbb{P}(s_i = x\mid s,\Omega, S,m)$). 
We define as a variant of the
 \emph{Inside-Outside algorithm}~\cite{Lari1990}, allowing us to obtain the
 desired probability,  the two  functions
$\Z{i,j}{m}{a,b}$ and $\Y{i,j}{m}{a,b}$. The former, as the 
equivalent of the \emph{inside}, computes the partition function between $i$ and $j$ included, 
knowing that position $i-1$ is composed of nucleotide $a$ (resp. $j+1$ is $b$) and containing $m$ mutations. The latter computes the \emph{outside}, in particular,  
the partition function considering only between $[0,i]\cup[j,n-1]$
knowing  that position $i+1$ is composed of $a$ (resp. $j-1$ is $b$) and containing 
 $m$ mutations outside.

The Boltzmann weights are a combination of the base pairs stacking energy, 
using as values those of the NNDB~\cite{Turner2010}, and
the average isostericity difference
between the mutant and $\Omega$, and $s$ with $\Omega$, using the isostericity values
 as defined in~\cite{Stombaugh2009}. The value of $10$ was used for the isostericity
 of any base pair 
 compared with $GG$ given that the latter base pair was not
  found in the~\cite{Stombaugh2009} tables,
 and the range of values is $[0,9.7]$, $0$ for a perfect isostericity.
